---
title: "Using bicPCA on Seurat objects"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Seurat-objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7, 
  fig.height = 5
)
```

We first load the BicPCA package

```{r setup}
library(BicPCA)
library(ggplot2)
library(Seurat)
library(patchwork)

seed = 777
set.seed(seed)
```

On this test, we use the spleen data from Tabula Muris, which provides a Seurat object. The data can be downloaded from this link: https://ndownloader.figshare.com/files/13092395 

```{r}
file <- "/Users/Data/"    #use your preferred directory
load(file = paste0(file,"facs_Spleen_seurat_tiss.Robj"))
x <- UpdateSeuratObject(tiss)
rm(tiss)
x
```

# Standard workflow

According with the elbow plot we would like to use between 6 to 20 dimensions, but this is a really subjective way to do it. Minka (2000) provides an alternative way using the Bayesian Information Criterion (BIC) over a probabilistic PCA. We then need to calculate the BIC for a desired set of PCs and define the number of PCs to use as those with the minimum BIC score.

```{r}
x <- NormalizeData(x, verbose = FALSE)
x <- FindVariableFeatures(x, verbose = FALSE)
x <- ScaleData(x, verbose = FALSE)
x
```

We can directly use our Seurat object on the FindNumberPC function to calculate the BIC for a set of PCs (e.g. using dim2Test = 40, the function will calculate the BIC for the set [1,40] of PCs). If we set returnBIC = TRUE, we will obtain the BIC scores for each of the tested dimension. 

```{r}
selDim <- FindNumberPC(x, dim2Test = 40, returnBIC = TRUE)
selDim$selPCA
plot(selDim$BIC)
```
We can use the selected dimensions on downstream analyses as visualization.

```{r}
x <- RunUMAP(x, reduction = "pca", dims = 1:selDim$selPCA, verbose = FALSE)
DimPlot(x, group.by = "cell_ontology_class", label = TRUE, label.size = 4) + NoLegend() + NoGrid()
```













